---
title: "peerass1.Rmd"
author: "YI HAN"
date: "Sunday, April 12, 2015"
output: html_document
---

# Loading and preprocessing the data
```{r, echo=TRUE}
setwd("C:/Users/vannem.han/OneDrive/2_Coursera/JHU_Repro_Res/PeerAssess/1")
OriginalData <- read.csv(file="activity.csv", head=TRUE, sep=",")
```

# What is mean total number of steps taken per day?
```{r, echo=TRUE}
Mydata <- OriginalData
Date <- Mydata[,2]
DateTable <- table(Date)

NumIntervalPerDay <- DateTable[1] #each day has 288 intervals  
NumDate <- length(DateTable) #It has 61 days

StepsPerDay <- rep(0, NumDate)

for(i in 1:NumDate)
  StepsPerDay[i] = sum(Mydata[(1+NumIntervalPerDay*(i-1)):(NumIntervalPerDay+NumIntervalPerDay*(i-1)),1])

StepsPerDay = StepsPerDay[!is.na(StepsPerDay)]

hist(StepsPerDay, breaks=30, xlab="Steps Per Day", main="Histogram of Steps Per Day", xlim=c(0,25000), ylim=c(0,12))
mean(StepsPerDay)
median(StepsPerDay)
```
2. The above pic is the histogram of the total number of steps taken each day. 
3. The mean and median of the otal number of steps taken per day is 10766.19 and 10765.

# What is the average daily activity pattern?
```{r, echo=TRUE}
StepsPerInterval <- rep(0, NumIntervalPerDay)

for(i in 1:NumIntervalPerDay)
  StepsPerInterval[i] = mean(Mydata[seq(from=i, to=i+NumIntervalPerDay*(NumDate-1), by=NumIntervalPerDay),1], na.rm=TRUE)

Interval <- Mydata[,3]
IntervalTable <- table(Interval)
IntervalNames <- names(IntervalTable) #interval names like 0 5 10 ... for plot

StepsPerInterval <- ts(StepsPerInterval)
plot(StepsPerInterval,xaxt='n', main="Time series of Averaged Steps Per Interval", ylab="Averaged Steps Per Interval", xlab="Interval")
axis(1, at=seq(1,NumIntervalPerDay,50),labels=IntervalNames[seq(1,NumIntervalPerDay,50)])
#max(StepsPerInterval)  
#which.max(StepsPerInterval)   #This-th 5 minute interval 
IntervalNames[which.max(StepsPerInterval)] #This 5 minute interval
```
1. The above pic is the time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.
2. The 835-th 5-minute interval, on average across all the days, contains the maximum number of steps.

# Imputing missing values
```{r, echo=TRUE}
Mydata <- OriginalData
sum(is.na(Mydata[,1])) #the total number of missing values
NAposition = which(is.na(Mydata[,1]))

for(i in NAposition)
  Mydata[i,1] = StepsPerInterval[which(IntervalNames==Mydata[i,3])]  #My strategy: the mean for that 5-minute interval

StepsPerDay_fillin <- rep(0, NumDate)

for(i in 1:NumDate)
  StepsPerDay_fillin[i] = sum(Mydata[(1+NumIntervalPerDay*(i-1)):(NumIntervalPerDay+NumIntervalPerDay*(i-1)),1])

hist(StepsPerDay_fillin, breaks=30, xlab="Steps Per Day (NA filled in)", main="Histogram of Steps Per Day (NA filled in)")
mean(StepsPerDay_fillin)
median(StepsPerDay_fillin)
```
1. 2304 is the total number of missing values in the dataset.
2. I use the mean for that 5-minute interval for filling in all of the missing values in the dataset.
4. The mean and median of the NA-filled dataset is 10766.19 and 10766.19. The mean value is the same from the previou one, while the median value changes. This is because I use the mean for that 5-minute interval to fill in the NA. 

#Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
Sys.setlocale("LC_TIME", "English") #set weekdays to be in English language

DateNames <- names(DateTable)

DayorEnd <- weekdays(as.Date(DateNames))
Seq_end <- c(which(DayorEnd=="Saturday"),which(DayorEnd=="Sunday"))

DayorEnd[Seq_end] = "weekend"
DayorEnd[which(DayorEnd!="weekend")] = "weekday"
DayorEnd = factor(DayorEnd)

Weekend <- rep(0, NumIntervalPerDay)
Weekday <- rep(0, NumIntervalPerDay)

for(i in 1:NumIntervalPerDay)
{ DateOneInterval <- seq(from=i, to=i+NumIntervalPerDay*(NumDate-1), by=NumIntervalPerDay)#coresponse to DateNames and DayorEnd.
  
  Weekend[i] = mean(Mydata[DateOneInterval[Seq_end],1])
  Weekday[i] = mean(Mydata[DateOneInterval[-Seq_end],1])
}

df <- data.frame(IntervalNames, Weekend, Weekday)
library(reshape2)
mm <- melt(df,id.var="IntervalNames")
library(ggplot2)
ggplot(mm, aes(IntervalNames, value, group=1)) + 
  #geom_point() + 
  geom_line() + 
  facet_grid(variable~.) + 
  labs(x="Interval",y="Number of steps",title="Steps Per Interval")
```

2. The above pic is the panel plot containing a time series plot of the 5 minutes interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
